---
title: "Machine Learning1"
author: "Tianyun Hua (PID:A91041691)"
date: "10/22/2021"
output: html_document
---

# Clustering Methods

Kmeans clustering in R is done with the 'kmeans()' function
Here we make up some data to test and learn with

```{r}
tmp <- c(rnorm(30, 3), rnorm(30, -3))
data <- cbind(x=tmp,y=rev(tmp))
plot(data)

```
Run 'kmeans()' set k(centers) to 2 (i.e. the number of clusters we want) nstart 20. The thing with Kmeans is you have to tell it how many clusters you want.

```{r}
km <- kmeans(data, centers = 2, nstar = 20)
km
```

>Q. How many points are in each cluster?

```{r}
km$size
```

>Q. What ‘component’ of your result object details cluster assignment/membership?
 
```{r}
km$cluster
```
 
 
>Q. What ‘component’ of your result object details cluster center?

```{r}
km$center
```
>Q.Plot x colored by the kmeans cluster assignment and add cluster centers as blue points

```{r}
plot(data, col = km$cluster)
points(km$centers, col = "blue", pch = 15, cex = 2)
```


# Hierarchail Clustering

We will use 'hclust()' function
```{r}
hc <- hclust(dist(data))
hc
```

hclust has a plot method
```{r}
plot(hc)
```
To find our membership vector we need to cut the tree and for this we use the 'cutree()' function and tell it the height to cut at. 

```{r}
cutree(hc, h = 7)
```

We can also use'cutree()' and state the number of clusters we want

```{r}
grps <- cutree(hc, k = 2)
plot(data, col = grps)
```

# Principal Componet Analysis (PCA)

## PCA of UK food data

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```

How many rows and cols?
```{r}
dim(x)
```

```{r}
rownames(x) <- x[,1]
x <- x[,1]
x
```

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names = 1)
x
```

```{r}
barplot(as.matrix(x), col = rainbow(17))
```

```{r}
barplot(as.matrix(x), col = rainbow(17), beside = TRUE)
```

```{r}
mycols <- rainbow(nrow(x))
pairs(x, col = mycols, pch = 16)
```

# PCA to my rescue

Here we will use the base R function for PCA, which is called 'prcomp()'. 
This function wants the transpose of data. 
```{r}

pca <- prcomp(t(x))
summary(pca)
```

```{r}
plot(pca)
```

We want score plot (PCA). Basically of PC1 vs PC2
```{r}
attributes(pca)
```

```{r}
plot(pca$x[,1:2])
text(pca$x[,1:2], labels = colnames(x))
```

We also examine the PCA "loadings", which tell us how much the original variables contribute to each new PC

```{r}
pca$rotation
barplot(pca$rotation[,1], las = 2)
```


## One more PCA for today
```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```

```{r}
ncol(rna.data)
colnames(rna.data)
```

```{r}
pca.rna <-prcomp(t(rna.data), scale = TRUE)
```

