---
output:
  pdf_document: default
  html_document: default
---
 ---
 title: 'calss 6 : R Functions'
 author: "Tianyun Hua PID:A91041691"
 date: "10/15/2021"
 output: github_document
---

## quick R Markdown intro
**bold** or *italic*

# This is a level 1 heading
### This is a level 3 heading


Here we will write a function to grade studetn homework.
New code chunck: option + command + I (for R code)

We will start with input example--student homework scores
```{r}
# example input vector
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```


> Q1. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

First, find the lowest score using **min()**, and the **which.min()** function to find the position of the smallest value

```{r}
student1
which.min(student1)

```

```{r}
student1[8]
# to get everything but the minimum value
student1[-which.min(student1)]
```

Then we can take the mean

```{r}
# this is my first solution
mean(student1[-which.min(student1)])
```

Is this a good idea?
```{r}
student2
# remove the NA value
mean(student2, na.rm = TRUE)
mean(student3, na.rm = TRUE)
# not a good idea
```

This is a bad idea because this is unfair using the na.rm = TRUE argument.

Let's change NA values to 0
Use the 'is.na()' function

```{r}
student2
is.na(student2)

x <- student2
x [is.na(x)] <- 0 
x
mean(x)
```

Combine working snippets to find the average for student3

```{r}
x3 <- student3
x3[is.na(x3)] <- 0
mean(x3[-which.min(x3)])
```

### now we can make our function

### use the "code"-"exrtact function"
```{r}

grade <- function(x) {
  x[is.na(x)] <- 0
  mean(x[-which.min(x)])
}
```

Now use the function to grade 3 students
```{r}
grade(student1)
grade(student2)
grade(student3)
```

Your final function should be
adquately explained with **code comments** and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” 
Now **grade the whole class**

Document the newly defined function with "code"-"Roxygen Skeleton"
```{r}
#' Calculate average scores for a vector of homework scores
#' dropping lowest single score. missing value will
#' be treated as zero score
#' @param x numeric vector of hw scores
#'
#' @return average score
#' @export
#'
#' @examples
#'  student <- c(100, NA, 90, 80)
#'  grade(student)
#' 
grade <- function(x) {
  # map NA missing homework values to zero
  # missing homework is 0 
  x[is.na(x)] <- 0
  # exclude lowest score hw
  mean(x[-which.min(x)])
}
```

```{r}
url <-"https://tinyurl.com/gradeinput"
gradebook<- read.csv(url, row.names = 1)
gradebook
```

We are going to use the super useful **apply()**

```{r}
apply(gradebook, 1, grade)
```

> Q2 Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook?

```{r}
results <- apply(gradebook, 1, grade)
# first approach
sort(results, decreasing = TRUE)
# second approach
which.max(results)
```


>Q3 From your analysis of the gradebook, **which homework** was **toughest** on students (i.e. obtained the lowest scores overall?

Here we want to calculate a summary stat for each column of the gradebook. Which stat should we use?

```{r}
# Let's try average, by column (margin = 2)
apply(gradebook, 2, mean)
```

 I can ignore the NA missing values
```{r}
hw.ave <- apply(gradebook, 2, mean, na.rm = TRUE)
which.min(hw.ave)
```

```{r}
hw.med <- apply(gradebook, 2, median, na.rm = TRUE)
which.min(hw.med)
```

Difference when using mean and median, plot the data and see

```{r}
boxplot(gradebook)
```

>Q4. Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)?

we will use **cor()** here

```{r}
gradebook[is.na(gradebook)] <- 0
cor(results, gradebook$hw5)
apply(gradebook, 2, cor, x = results)
```

Homework 5 is the most predicative of overall score

